---
title: "Soil respiration variability and correlation across a wide range of temporal scales"
author: "Ben Bond-Lamberty"
date: "28 October 2019"
output: html_document
---

## Introduction

* Noodling


```{r setup, include=FALSE}
library(drake)
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(kableExtra)
library(lattice)
library(cowplot)  # 0.9.4
theme_set(theme_bw())
#theme_set(theme_cowplot())

N_HISTOGRAM_BINS <- 20
START_DATE <- "2017-10-01"
END_DATE <- "2099-08-01"

readd("con_licor_data") %>% 
  filter(Timestamp >= START_DATE, Timestamp < END_DATE) %>% 
  arrange(Timestamp, Port) ->
  cld

message("Data limits: ", START_DATE, " to ", END_DATE)
```


## Methods

```{r filtering, echo=FALSE}
cld %>% 
  filter(R2 > 0.75, !is.na(Flux)) %>%  # probably a chamber closing problem otherwise
  group_by(week(Timestamp)) %>% 
  mutate(mad = abs(Flux - median(Flux,  na.rm = TRUE)) / mad(Flux, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(mad <= 4) %>% 
  ungroup() ->
  cld_clean
removed_cld <- nrow(cld) - nrow(cld_clean)
```

Continuous data: `r nrow (cld)` observations; `r removed_cld` (`r round(removed_cld / nrow(cld) * 100, 0)`%) removed because of chamber closure or other problems. Final data is `r nrow(cld_clean)` observations.

Have data for `r length(unique(yday(cld_clean$Timestamp)))` unique days of the year.

```{r missing-dates, echo=FALSE}
# Insert missing dates
cld_clean %>% 
  mutate(Year = year(Timestamp), Month = month(Timestamp),
         Yday = yday(Timestamp), Hour = hour(Timestamp)) %>% 
  # complete() will fill in a few impossible dates, generating warnings below
  complete(Year, Month = 1:12, Yday = 1:365, Hour = 0:23, Port = 1:8) %>% 
  mutate(Timestamp = if_else(is.na(Timestamp), 
                             suppressWarnings(ymd_h(paste(Year, Month, Yday, Hour))),
                             Timestamp)) %>%
  # complete() also adds 'missing' data before and after we measured; remove
  filter(!is.na(Timestamp),
         Timestamp >= min(cld_clean$Timestamp),
         Timestamp <= max(cld_clean$Timestamp)) %>% 
  select(Timestamp, Year, Month, Yday, Hour, Port, Flux) ->
  cld_clean_hourly
```


```{r continuous-plot, echo=FALSE,fig.height=8}
cld_clean_hourly %>% 
  mutate(Year = as.factor(year(Timestamp))) %>% 
  ggplot(aes(yday(Timestamp), Flux, color = Year)) + 
  geom_line(alpha = 0.75) + facet_grid(Port ~ .) +
  ggtitle("A. Continuous data")
```


```{r continuous-collar-cv, echo=FALSE}
cld_clean_hourly %>% 
  mutate(Year = as.factor(year(Timestamp)),
         Hour = hour(Timestamp)) %>% 
  group_by(Year, Yday, Hour) %>% 
  summarise(Flux_cv = sd(Flux) / mean(Flux)) %>% 
  ungroup() ->
  cld_cv

cld_cv %>% 
  ggplot(aes(Yday, Flux_cv, color = Year)) + 
  geom_line() +
  ggtitle("B. Collar-to-collar CV")

cld_clean_hourly %>% 
  mutate(Month = month(Timestamp),
         Hour = hour(Timestamp)) %>% 
  group_by(Month, Hour) %>% 
  summarise(Flux_cv = sd(Flux, na.rm = TRUE) / mean(Flux, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(Hour, Flux_cv, color = Month, group = Month)) + 
  geom_line() +
  ggtitle("C. Collar-to-collar CV by month")
```

As we changing sampling frequency, how does our estimate of Rs_annual change?

```{r monte-carlo, cache=TRUE}
cld_clean_hourly %>% 
  arrange(Timestamp) %>% 
  group_by(Port) %>% 
  mutate(Flux_approx = approx(Timestamp, Flux, xout = Timestamp)$y) %>% 
  filter(!is.na(Flux_approx)) ->
  cld_clean_gapfill


compute_frac_mean <- function(frac, df) {
  df %>% 
    sample_frac(frac) %>% 
    pull(Flux_approx) %>% 
    mean()
}

# This isn't what we want
# Want to compute the time length of the series; sample some fraction;
# and then adjust to an annual (or monthly or whatever) rate
# We want the "sample per year" variable to go to 365 * 24

set.seed(12345)
N_MC <- 100
cld_clean_gapfill %>% 
  group_by(Port) %>% 
  summarise(meanflux = mean(Flux_approx)) ->
  meanflux

results <- list()
for(p in unique(cld_clean_hourly$Port)) {
  x <- filter(cld_clean_gapfill, Port == p)
  results[[p]] <- tibble(Port = p,
                         frac = rep(seq(0.01, 1, by = 0.05), times = N_MC),
                         N_per_month = 365.25 * frac / 12,
                         flux = sapply(frac, compute_frac_mean, x))
}
bind_rows(results) %>% 
  left_join(meanflux, by = "Port") %>% 
  mutate(flux_error = (flux - meanflux) / meanflux) %>% 
  group_by(Port, N_per_month, frac) %>% 
  summarise(n = n(), 
            flux_mean = mean(flux), 
            flux_sd = sd(flux),
            flux_error_mean = mean(flux_error),
            flux_error_sd = sd(flux_error)) ->
  y

ggplot(y, aes(N_per_month, flux_error_mean)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = flux_error_mean - flux_error_sd, 
                  ymax = flux_error_mean + flux_error_sd), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = c(1, 4, 10, 20)) +
  coord_trans(x = "log10") + 
  facet_wrap(~Port)
```



## Session

```{r, echo=FALSE}
sessionInfo()
```
