---
author: "SPennington"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Diagnostics as of:
```{r, echo = FALSE} 
date()
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
figures <- get(load("../outputs/figures.rda"))
plots <- read.csv("../design/plots.csv")
knitr::kable(plots) 
```

|Location |Experiment           |No. of Collars |
|:-------:|:-------------------:|:-------------:|
|SERC     |Control              |56             |
|SERC     |Salinity transplant  |36             |
|SERC     |Elevation transplant |32             |

### Flux over time, color is indicating plot each core originated from
(left to right - High, Medium, Low salinity); (top to bottom - Low, Medium, and High elevation)
```{r, echo = FALSE}
figures$timeflux_plot
```

### Composite temperature and flux data for all cores
```{r, echo = FALSE, warning = FALSE}
figures$q10_plot
```

### CV distribution for each collar, n = number of measurements taken
```{r, echo = FALSE}
figures$ggCV
```